### 2020-08-24 code/figure2.Rmd

```{r figure-2-plot, fig.width = 9}
load("../data/figure1B_data.rda") ## Load in data processed in figure1.Rmd
load("../data/optimal_minpts_score.rda")
umap_cancertype <- data.table::fread("../data/umap_3d_coors.tsv")
load("../data/kelly.colours.rda")
load("../data/ditto.colours.rda")

opar <- par(mfrow = c(2,3), mai = c(0.1, 0.1, 0.1, 0.1)) ## plot layout 2x3
par(mar = c(4,6,2,0) + 0.1) ## setting plot margins for the following plots
## Figure 2A: Explained variance of eigenvalues
with(randomize %>% filter(eigenvalue < 50), plot(x = eigenvalue, y = expl_var / 100, ## plotting the explained variance of observed
                                                 type = "l", yaxt = "n", bty = "l",
                                                 col = "#998ec3", 
                                                 xlab = "", ylab = ""))
with(randomize %>% filter(eigenvalue < 50), lines(x = eigenvalue, y = perm / 100, ## plotting the explained variance of permuted
                                                  col = "#f1a340"))
yticks_val <- scales::breaks_pretty(n=5)(randomize %>% filter(eigenvalue < 50) %>% pull(expl_var) / 100)
axis(2, at=yticks_val, lab=scales::percent(yticks_val), las = 1) ## Axis labels
title(ylab = "explained variance", line = 3.5)
title(xlab = "MCA Eigenvalues", line = 2)
legend("topright", legend=c("Eigenvalues", "Permuted Eigenvalues"),
       col=c("#998ec3", "#f1a340"), lty=1, cex=0.8)
mtext("A", side=3, line=0, cex=1.2, adj=-0.35) ## plot tag

par(mar = c(2,2,0,0), bg = NA) ## setting plot margins for the following plots
## Figure 2C: MCA projected down to 3 dimensions (UMAP)
with(umap_cancertype, scatter3D(plot_y, -plot_z, plot_x, 
                                bg = "white", pch = 21, cex = 0.8, lwd = 0.2,
                                theta = 0, phi = 65, scale = F,
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
title(xlab="UMAP 2", line=0)
mtext("C", side=3, line=-2.05, cex=1.2, adj=0) ## plot tag

par(mar = c(4,6,2,0) + 0.1) ## setting plot margins for the following plots
## Figure 2D: Optimization of HDBscan
with(sample_df, plot(x = minPts, y = value, 
                     col = "grey", pch = 20, yaxt = "n",
                     xlab = "", ylab = "", bty = "l",
                     las = 1, cex = 0.3))
with(score_df, lines(x = minPts, y = score,
                     col = "black", las = 1))
abline(v = optimal_minPts, col = "red") ## Red vertical line to show optimal minimum point for HDBscan
title(ylab = "score", line = 3.5)
title(xlab = "Minimum Cluster Size", line = 2)
axis(2, at=c(0.25, 0.35, 0.45, 0.55), las = 1)
mtext("D", side=3, line=0, cex=1.2, adj=-0.27) ## plot tag

## Figure 2B: pvalue of eigenvalue 
with(randomize %>% filter(eigenvalue < 50), plot(x = eigenvalue, y = pvalue,  ## line plot
                                                 type = "l", bty = "l",
                                                 col = "black", las = 1,
                                                 xlab = "", ylab = ""))
with(randomize %>% filter(eigenvalue < 50), points(x = eigenvalue, y = pvalue, pch = 20, ## add dots to line plot
                                                   col = "black"))

## Determining the optimal mca dimension
optimal_mca_dimensions <- randomize %>%
  filter(pvalue < 0.05) %>% ## filtering for pvalue < 0.05
  filter(eigenvalue == max(eigenvalue)) %>% ## filtering for the maximum eigenvalue
  pull(eigenvalue)

abline(v = optimal_mca_dimensions, col = "red") ## Red vertical line to show optimal MCA dimension
title(main = paste0("Significant dimensions = ", optimal_mca_dimensions), adj = 0, font = 3, cex.main = .8) ## Axis labels
title(ylab = "P-value", line = 3.5)
title(xlab = "MCA Eigenvalues", line = 2)
mtext("B", side=3, line=0, cex=1.2, adj=-0.35) ## plot tag

par(mar = c(2,2,0,0), bg = NA) ## setting plot margins for the following plots
## Figure 2E: HDBscan
cluster.colours <- c("grey", kelly.colours[c(3:12)]) ## Color scheme for cluster
names(cluster.colours) <- c(0:10) ## assigning cluster to the colors
with(umap_cancertype, scatter3D(plot_y, -plot_z, plot_x, 
                                bg = cluster.colours[as.character(clust)], pch = 21, cex = 0.8, lwd = 0.2, # clust is hdbscan result
                                theta = 0, phi = 65, scale = F,
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
title(xlab="UMAP 2", line=0)
mtext("E", side=3, line=-2.05, cex=1.2, adj=0) ## plot tag

## Figure 2F: kNN 
with(umap_cancertype, scatter3D(plot_y, -plot_z, plot_x, 
                                bg = cluster.colours[as.character(clust_knn)], pch = 21, cex = 0.8, lwd = 0.2, # clust_knn is knn result
                                theta = 0, phi = 65, scale = F,
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
title(xlab="UMAP 2", line=0)
mtext("F", side=3, line=-2.05, cex=1.2, adj=0) ## plot tag

>>>>

```{r figure-2-plot, fig.width = 9}
umap_cancertype <- data.table::fread("../data/umap_3d_coors.tsv") ## load data
load("../data/kelly.colours.rda")
cluster.colours <- c("grey", kelly.colours[c(3:12)]) ## Color scheme for cluster
names(cluster.colours) <- c(0:10) ## assigning cluster to the colors

theta <- c(0, 90, 270)
phi <- c(65, 15, 15)

opar <- par(mfrow = c(1,3), mai = c(0.1, 0.1, 0.1, 0.1)) ## plot layout 2x3
par(mar = c(2,2,0,0), bg = NA) ## setting plot margins for the following plots

for(n in 1) {
  with(umap_cancertype, scatter3D(plot_y, -plot_z, plot_x, 
                                  bg = cluster.colours[as.character(clust_knn)], pch = 21, cex = 0.8, lwd = 0.2, # clust_knn is knn result
                                  theta = theta[n], phi = phi[n], scale = T,
                                  xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                  colvar = NULL))
  title(xlab="UMAP 2", line=-5.2)
  for(p in 2:length(phi)) {
    with(umap_cancertype, scatter3D(plot_y, -plot_z, plot_x, 
                                bg = cluster.colours[as.character(clust_knn)], pch = 21, cex = 0.8, lwd = 0.2, 
                                theta = theta[p], phi = phi[p], scale = T,
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
  }
}
